---
title: "Bonza Snowshoe Exploratory Report"
author: "Grace Bianchi"
date: "11/6/2021"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(effsize)
library(lubridate)
library(ggbeeswarm)
```

### 1. Overview
This report provides an exploratory overview 

Introduction: A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.

### 2. Data and methods
A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

### 3. Prelimnary results

#### A. Annual juvenile hare trap counts
```{r}
# read in the data
snowshoe_juveniles <- read_csv(here("data", "bonanza_hares.csv")) %>% 
  # convert date to date class and put output in new column
  mutate(date = mdy(date)) %>% 
  # add new column with just the year 
  mutate(year = year(date)) %>% 
  # filter for juveniles
  filter(age == "j") %>% 
  mutate(sex = case_when(
    sex == "f" ~ "female",
    sex == "m" ~ "male"))
  
```

```{r}
snowshoe_counts <- snowshoe_data %>% 
  # group by year
  group_by(year) %>% 
  # use summarize to get counts for each year
  summarize(counts = n())

ggplot(data = snowshoe_juveniles, aes(x = year)) +
  geom_histogram(bins = 15, color = "black") +
  scale_x_continuous(breaks = c(1998, 2000, 2002, 
                                2004, 2006, 2008, 2010, 2012)) +
  labs(title = "Annual juvenile hare trap counts",
       x = "Year",
       y = "Count")
```

Figure 1. Annual juvenile hare trap counts 

What were the minimum and maximum counts of juvenile hare trappings?
● Mean or median annual number of juvenile hares trapped?
● General/notable trends?
● These are absolute counts, but this count is also impacted by
effort (e.g. the number of days and number of traps in the study each year). What would you suggest be done moving forward (you are not expected to actually do this) to standardize the juvenile hare population in future analyses?

#### B. Juvenile hare weights
```{r}
ggplot(data = snowshoe_juveniles, aes(y = weight)) +
  geom_histogram(aes(fill = sex), bins = 15, position = "dodge") +
  facet_grid(rows = vars(grid))
```

#### C. Juvenile weight comparison by sex

```{r}
juvenile_weight_sex <- snowshoe_juveniles %>% 
  group_by(sex) %>% 
  summarize(mean = mean(weight, na.rm = TRUE),
            sd = sd(weight, na.rm = TRUE),
            sample_size = n())

juvenile_weight_sex %>% 
  kable(col.names = c("Sex",
                      "Mean",
                      "Standard Deviation",
                      "Sample Size"),
        digits = 2,
        caption = "Juvenile Weights by Sex") %>% 
  kable_styling(full_width = FALSE)

```

A 2 - 3 sentence comprehensive statement of the means comparison, including at least:
○ The actual difference in means
○ Effect size (Cohen's *d*) (see effsize::cohen.d() example in
the Lab 5 key)
○ Outcome of an unpaired two-sample t-test

```{r}
sex_cohen <- cohen.d(female, male)
```


Data citation: Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14
